Global adjoint tomography -- Model GLAD-M25

### 2018-10-09 main.tex
Updates from ShareLaTeX

\begin{figure}
  \includegraphics[width=0.8\textwidth]{figures/station_map.pdf}
  \caption{Distribution of 11,800 seismographic stations on the globe. Colors denote the number of events which they contribute waveforms to in the model inversion stage. Stations with event response < 400 are plotted in smaller size and they are usually temporary arrays which are deployed over a short period of time. The max event reponse comes from IU.ANMO, which contributeed to 1442 out of 1480 eartquakes in out dataset.}
  \label{fig:stations}
  \centering
\end{figure}

@@

The misfit for source~$s$, category~$c$, receiver~$r$, and window~$w$ is
\begin{align}
  \chi_{scrw} = \int \Big[ \frac {\Delta \tau_{scrw}} {\sigma_{scrw}} \Big]^2 d\omega
\quad ,
\end{align}
is a frequency-dependent phase measurements using then multitaper technique.
where $\Delta m_{scrw}$ denotes a measurement with associated standard deviation~$\sigma_{scrw}$.-

@@

\subsection{Model Parametrization}
Transversely isotropy model parametrization is used in our model, same as S362ANI(GLAD-M00)
and GLAD-M15. Comparing to a general anisotropic model which has 21 independent variables,
our model could be described by 5 Love parameters, A, C, L, N and F\citep{love2013treatise}.
Since it is more straight forward to use  wavespeed as model parameters in sesimology,
we choose $\alpha_v$, $\alpha_h$, $\beta_v$, $\beta_h$ and $\eta$.
Assuming the radial anisotropy is due to shear anisotropy, these 5 parameters
can be further reduced to 4, we introduced the bulk sound speed,
$c=\sqrt{\kappa/\rho}$. Our final parameters are:\\
\begin{align*}
      c &= \sqrt{\kappa/\rho} \\
\beta_v &= \sqrt{L/\rho} \\
\beta_h &= \sqrt{N/\rho} \\
\eta & = F/(A-2L)
\end{align*}

@@

where $\beta$ is defined as the Voigt average of the raidally anisotropic shear wave speeds\citep{babuska1991seismic}
$$\beta = \sqrt{(2\beta_v^2 + \beta_h)/3}$$

@@

The flow chart for adjoint tomography is shown in Fig.\ref{fig:adjoint_workflow}.
The workflow starts with a certain number of earthquakes. Here, we used the
inverted CMT sources from the Global CMT catalog. Based on the time
of earthquakes, we downloaded observed data, including time series and StaionXML,
from IRIS and other data centers. The we convert the sources, StationXML and time series
into ASDF files. Each ASDF file usually contains a complete set of data from one
earthquake. The data download and conversion need to be done only once. The raw observed data,
storing in ASDF, could be reused afterwards.

@@

To simulate 120min of seismograms at 17sec resolution, it takes 384 Tesla K20X
GPUs approximately 10 mins to run on Titan. To precisly calculate the sensitivity
kernels, there is forward wavefield saved to the disk and those will used in the
adjoint simluation in the later stage. In the current resolution
and simulation length, the forward wavefield saved is above 1 terabytes
for each earthquake. Given the size the 1480 earthquakes, we have 1.5 petabytes of
wavefield saved to disk using ADIOS\citep{liu2014hello} during about 10 hours
of simulation using 15,360 GPUs.
It is huge volume of data saved to distk in a relatively short period of time, in which case
ADIOS really helped to relief the I/O bottleneck.

@@

The seismic data processing takes the raw observed and synthetic data as input,
and generate s the adjoint sources as the output. As shown
in Fig.\ref{fig:preprocess_workflow}, preprocessing includes:
\begin{enumerate}
  \item Signal processing that remove the instrument response from observsed data
    to recover the true ground motion. Both observed and synthetic data will be
    filtered into a certain period band. Then the two horizontal components of the
    sesimograms are roated from North, East to radial and transverse only if the
    StationXML passed the sanity check.
  \item Window selection that picks the time window on a pair of
    observed and synthetic seismograms. Pyflex(python version
    of FLEXWIN\citep{maggi2009automated}), is
    such a tool that automatically generate windows where observed data and
    synthetic data are close enough to make measurements based on user defined
    criteria.
  \item Multi-taper measurement that generates frequency-dependent travel-time
    difference within the windows selected from the last step.
    The output files contains both the measurements and misfit values.
  \item Window cleaning that get rid of bad windows(and associated
    measurements), only keeping good ones based on the statistical analysis of
    measurements.
  \item calculate the adjoint sources with the sorted windows.
  \item For each window and associated measurement, calculate its weightings
    that will be applied at the summation stage and calibrate its contribution
    into the overall misfit function, using the methond we mentioned before.
  \item Construct the final adjoint sources by summing the adjoint source from
    different windows and measurement categories.
\end{enumerate}

Even though it is such a computational demanding process as the forward and adjoint
simulation, only taking about $1\%$ CPU hours in the whole workflow, it is crucial
that directly impact the inversion results. It decides which measurement
will be taken into consideration and have a footprint in the final gradient.
We are very careful at this stage, including a lot of checks, such as the Station and
response integratity check, and also filtering out bad measurements to avoid big outliers.
On the other hand, it is very process-demanding, which means it has complex data and
control flows, that every seismograms will go through quite a few of processes
before the final results is generated. We have millions of seismic traces and ten
millions of windows, it is impossible to human-check every trace and measurements.
I will go into more depth about this part in the following sections, discussing our
efforts how to make the data and process safe and robust.

@@

The post-processing takes the frechet derivatives as the input and generate
the model updates as the output. It contains:
\begin{enumerate}
  \item Summation of each kernel for each earthquake to obtain one
    gradient for the overall misfit function. The number of kernel
    files will reduce from 1480 to 1 after summing them up. Also,
    there will source weightings applied at this stage, to balance
    the uneven distribution of earthquakes.
  \item Smoothing the misfit gradient using 3D gaussian function, which
    serves as a regularization procedure. Instead of using a changing
    smooth radius based upon the "ray density"\citep{bozdaug2016global},
    we used the fixed value in the same iteration, following\cite{zhu2012structure}.
  \item Precondition the smoothed kernels. Following
    \citep{luo2013strategies, zhu2012structure, bozdaug2016global},
    the preconditioner we used is:
    \begin{equation}
      P(\mathbf{x}) = 1 / \int \partial_t^2 \mathbf{s}(\mathbf{x}, t) \cdot \partial_t^2 \mathbf{s}^\dagger(\mathbf{x}, T-t) dt
    \end{equation}
\end{enumerate}

@@

Based upon the misfit gradient obtained from the post-processing stage.
Line search is then used to determine the step length for model update
along the search direction. Forward simulations were run for a subset
of 120 earthquakes with different step length and the one with
the max misfit reudction was picked as the model update step length.

@@

The Adaptable Seismic Data Format, ASDF, serves as a self-contained-and-explained data container while taking full advantages of parallel computing. In our case, one ASDF file contained all the files need in data processing, including the seismic traces, quakeml files and station response files. The APIs were carefully designed for easy data extraction and parallel implementation. Each trace in the ASDF is associated with one specific quakeml file and stationXML file, reducing the possibilities of making mistakes.

@@

The inversion went through a few stages, mainly due to adding more data
and measurement categories and fine tunning the weighting strategies.

@@

The major change at stage II is we split the 40-100 sec period band into two: one with
body wave and the other with only surface wave measurements. Since body waves and
surfaces are sensitive to the different parts of the earth, by splitting these two
phases, we have more controls over spatical distribution of the model update.
Table.\ref{table:measurement_category} shows the measurement categories we used starting
from M18 and we will stick to it till M25.

@@

Since surface waves are mostly sensitive to shallow
structures and for shallow structures our model performs quite good
already given current dataset, we don't expect to see large misfit
reduction from in those long period surface wave category.
To better bring out the deep structure, We tune
down the weightings of 40-100 sec surface wave, thus reduce their
contribution in the overal misfit function. That explained why we see the
misfit jum(decrease) from M17 to M18 in Fig.\ref{fig:misfit}(h) - (j) and
increase Fig.\ref{fig:misfit}(b)-(d). We are using the same data but
redistribution the contribution from each category by changing the measurements
weight in the overall misfit function.

@@

The misfit values for 17-40sec and 40-100s body wavess are decreasing fast at
the first few iterations, and flattens gradually, which may indicate that we are
approaching the model resolution limits given current dataset.

@@

We added an extra set of 440 earthquakes into our inversion database. Thus, the total
number of earthquakes increased from 1,040 to 1480. We also bring back weightings
of 40-100sec surface wave as stage I since the newly added 440 was mostly located
in poorly covered regions in our previous dataset and those surface wave meansurements
will help to better constrain the shallow structures at those regions.

The misit jumps, from M21 to M22, is mostly due to the redistribution of weights
and new earthquakes being added to the dataset. Optimization method is changed
from conjugate gradient to L-BFGS. Overall misift values is steadily reducing,
as well as those in each category.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/misfit.pdf}
  \caption{Evolution of phase misfits from GLAD-M15 to GLAD-M25.
  Each color denotes the different stages of inversions, between
  which we make changes to the inversion by either tuning the
  weighting strategies or adding more data.
  (a) Evolution of the total phase misfit. (b) - (j) Misfit in
  each period band and component. (e) - (g) From GLAD-M18,
  we future split the 40-100 sec period band into two
  measurement categories: body(red dimond) waves and surface waves
  (red square).}
  \label{fig:misfit}
\end{figure}

@@

In the last setion, we discussed how the misfit evolves in iterations. However, since
we changed our dataset and misfit functions(mainly weightings) between different stages,
it might not be clear that how much did we improve our model. To precisely measure
our model performance, we calculate synthetic data using the same source using M00, M15
and M25. Windows and weightings are also identical across different iterations.

Table.\ref{table:misfit_reduction_M15_M25} shows the misfit reduction from M15 to M25
for 12 categories. For all categories, there is a significant misfit reduction. Among
the three components, transverse have the largest misfit recution. For the 4 period bands,
40-100s surface waves have the largest misfit reduction which may indicate the 40-100 sec
surface wave might not have good fit between synthetic and observed data compared to
other period band and our inversion is improving it a lot.

Table.\ref{table:misfit_reduction_M00_M25} shows M00 to M25. It is very similar to
Table.\ref{table:misfit_reduction_M15_M25} but with larger values since it also
counts the model improvements from M00 to M15. The best improved category is
the transverse component in 40-100s surface waves, with misfit reduction at
53.8\%. The minimum misfit reudction, on the other hand, is the vertical component
in 17-40s body waves, with 31.1\% misfit reduction.

@@

Another way to evaluate the model performance by checking the distribution
of measurements. Fig.\ref{fig:phase_hist} shows the histograms of phase
measurements. We use identical windows to make measurements using 1,480
earthquakes on the all the three models, our starting model
GLAD-M00(362ANI+CRUST2.0), first generation model GLAD-M15,
and lastest model GLAD-M25. The total number of selected windows is over
18 million. The standard deviations values of phase measurements has
significant improvments on all the categories over iterations.
The means values, has also improved expecially in the long period categories,
such as the 40-100 surface waves and 90-250 all waves. In some categories,
you may see slight shift of mean values, mainly due to the trade off among
different compnents and period bands. To summarize, the improvements in
the phase measurements histogram is consistent as we observed in the
misfit reduction.

@@

The other prominent feature of our models is that we can also constrain $V_p$ as an
independent model parameters. Fig.\ref{fig:global-vp} shows our isotropic $V_p$
structures at serveal depths. As comparisons, we also showed two more global
$V_p$ models, LLNL-G3Dv3\citep{simmons2012llnl}} and GAP-P4\citep{fukao2013subducted}.

@@

For upper mantle, the LLNL-G3Dv3 shows the largest variations. For example,
the 500km, the LLNL model shows very strong slow velocity anomalies at
Hawaii and other pacific hotspots, such as Samoa and Tahiti.
For middle mantle at about 1000km, all them models all shows a few major
subduction zones, such as the south america, Fiji and Sumatra regions.
At lower mantle, such as 2000km and 2800km, LLNL-G3Dv3 and GAP-P4 tends to
fade out, even though the patterns are very similar among them, such as
the LLSVP regions.

** Note: add sphereical hormonica analysis later **

@@

\subsubsection{North America}
The deployment of USArray has dramatically incresed the data overage on the
continental United States. In this section, we compare our GLAD-M25 with
together with two global models, S362ANI+M, S40RTS, and two regional models,
US22\citep{zhu2017radial}, SL2013NA\citep{schaeffer2014imaging}.

US22 is radially anosotropic model build based on adjoint tomography using
USArray data from 180 regional earthquakes using 15-50sec short-period
body waves and 25-150sec long-period surface waves.
SL2013NA is an uppe-matnle global shear wave speed model using multimode
surface wavefrom tomography and primally focused on the north american
region.

The North American region is featured by a large, high-velocity cartonic
lithosphere, bounded by the Rocky Mountain Front to the west and continental
margein to the east. For both sides the lithosphere has been deformed by
low velocity structures thus form a sharp boundary.

At 100km, our model shows a much sharpened boundary between the carton
and surrounding low veclocity strcutures, especially compared with
two global models, S362ANI+M and S40RTS, yet stay very similar to
two regional models, US22 and SL2014NA. At such depth,
we are observed a very strong expression of the Snake River Plain(
with YellowStone) and Raton hotspot, with their depth extending to 200km.
The very high-velocity caraton extend to 200km, with its very core part to
300 km, similiar to US22 while SL2013NA shows it almost disappeared at such depth.
Again, the Wyonming craton and Mecicine Hat Block shows a very strong lithospheric
root from the very shallow depth to 400km, which is also smiliar to US22 but
hard to observe in SL2013NA. Similiar case is the craton at Yavapai and
Mazatzal Blocks, which define the souther bound of the large craton.

At such depth, we also observed high velocity
features within the Gulf of Mexico, which conincides spatially with the deepest
bathymetry and corresponds to a portion of anciet oceanic
Lithosphere\citep{muller2008}. The remanant of Juan de Fuca subudction
plates, with its high velocity footprint, is presented in our model at
300km and 400km, though the SL2013NA shows a much stronger anomaly.

At the North, our models shows a extended high veloicty craton beneath the 
Greenland, though clearly seperated with the North America Craton,
bounded at the relativly low velcoity strcture at Barffin Bay and Labrador
Sea, with North American Craton followed nicely with the coastline. In the
north west region, our models shows the most clear picture of
Aleutian subduction zones, with it high velocity feature down to 400km.
In the nort easter part, the Iceland hotspot has been greatly enhanced,
especially at 300km and 400km compared to S362ANI+M.

In the Easter part, our models shows a high-velocity structure at the western
Atlantic, starting from the tip of Florica to the New York,
seperated from the major part the north america craton,
though it stays very thin that the depth only extend to above 200km.
This feature is associated with some of the oldest oceanic crust and lithosphere
within the Atlantic\citep{muller2008, schaeffer2014imaging}.
All the models is showing a high velocity anomaly at Nova Scotia and Newfoundland
with its high-velocity core to 300km.
Bermuda hotspot is nicely perserved from our starting
model, yet get sharpened and stronger at 200km and 300km. Besides, the
Carribean suduction, which is relatively young subduction systems, is clearly
observable in our model, starting from 200km to 300km.

@@

\subsubsection{Europe}
Fig.\ref{fig:europe-vs} shows the isotropic $V_s$ structures at Europe. As
comparisons, we also showed S362ANI+M, S40RTS and one regional model EU60
\citep{zhu2015seismic}. Compared to two global models, it is clear that
our model is much sharper at few locations. At 100 and 200km, Eifel
hotspot is clearly seen in our models, as well as in EU60. At 400km
and 600km, the subduction is very consistent between GLAD-M25 and EU60,
with the curved shape at the lower bottom of the figure,
while S362ANI+M shows very smooth shape and S40RTS shows much smaller 
perturbation values. Besides, at 400km, GLAD-M25 is enhancing Iceland
compared to S362ANI+M.

@@

We presented our model GLAD-M25, with S362ANI+M, S40RTS and EARA2014.
At 100km depth, our model shows sharpening at the Himalayas subductions
zones. At 250km, our models shows the subudction in the east part.
At 400km, our models shows the subduction at sumartra.

@@

\subsubsection{South America}

@@

\subsection{Held-out Dataset Test}
Following the approch by \cite{tape2009adjoint}, \cite{bozdaug2016global},
unseen events are used to check the performance of our
model. We revisited the Global CMT catalogue,
picked all the events whose magnitude is between 6.3 and 7.0, which hasn't
been used in our inversion. The reason why we set the lower bound to 6.3
instead of 5.5, is those earthquakes have good enough data and generates
more measurements at the global scale.
We calibrated the scalar moment and origin time in the GLAD-M25, and there
are 360 events left that has enough windows. Fig.\ref{fig:events_360} shows
the ditribution of 360 earthquake that we used as held-out dataset

@@

First, we evaluate the model performance by misfit reduction. Same windows and
weightings are used across models. Table.\ref{table:misfit_reduction_M00_M25_360}
shows the misfit reduction from M15 to M25 and Table.\ref{table:misfit_reduction_M00_M25_360}
is from M00 to M25. It agains shows misfit reduction in all categories in
our held-out dataset, simliar to what we observed from the inversion dataset.
The reduction are slightly smaller, partly due to the reason that we didn't
do waveform-type inversion as we mentioned in the source inversion stage but
only did a grid search to calibrate the scalar moment and origin time.

>>>>

\begin{figure}
  \includegraphics[width=0.8\textwidth]{figures/station_map.pdf}
  \caption{Distribution of 11,800 seismographic stations on the globe. Colors denote the number of events which they contribute waveforms to in the model inversion stage. Stations with event response < 400 are plotted in smaller size and they are usually temporary arrays which are deployed over a short period of time. The max event response comes from IU.ANMO, which contribute ed to 1442 out of 1480 earthquakes in out dataset.}
  \label{fig:stations}
  \centering
\end{figure}

@@

The misfit for source~$s$, category~$c$, receiver~$r$, and window~$w$ is
\begin{align}
  \chi_{scrw} = \int \Big[ \frac {\Delta \tau_{scrw}} {\sigma_{scrw}} \Big]^2 d\omega
\quad ,
\end{align}
is a frequency-dependent phase measurements using then multi-taper technique.
where $\Delta m_{scrw}$ denotes a measurement with associated standard deviation~$\sigma_{scrw}$.-

@@

\subsection{Model Parametrization}
Transversely isotropy model parametrization is used in our model, same as S362ANI(GLAD-M00)
and GLAD-M15. Comparing to a general anisotropic model which has 21 independent variables,
our model could be described by 5 Love parameters, A, C, L, N and F\citep{love2013treatise}.
Since it is more straight forward to use  wavespeed as model parameters in seismology,
we choose $\alpha_v$, $\alpha_h$, $\beta_v$, $\beta_h$ and $\eta$.
Assuming the radial anisotropy is due to shear anisotropy, these 5 parameters
can be further reduced to 4, we introduced the bulk sound speed,
$c=\sqrt{\kappa/\rho}$. Our final parameters are:\\
\begin{align*}
      c &= \sqrt{\kappa/\rho} \\
\beta_v &= \sqrt{L/\rho} \\
\beta_h &= \sqrt{N/\rho} \\
\eta & = F/(A-2L)
\end{align*}

@@

where $\beta$ is defined as the Voigt average of the radially anisotropic shear wave speeds\citep{babuska1991seismic}
$$\beta = \sqrt{(2\beta_v^2 + \beta_h)/3}$$

@@

The flow chart for adjoint tomography is shown in Fig.\ref{fig:adjoint_workflow}.
The workflow starts with a certain number of earthquakes. Here, we used the
inverted CMT sources from the Global CMT catalog. Based on the time
of earthquakes, we downloaded observed data, including time series and StationXML,
from IRIS and other data centers. The we convert the sources, StationXML and time series
into ASDF files. Each ASDF file usually contains a complete set of data from one
earthquake. The data download and conversion need to be done only once. The raw observed data,
storing in ASDF, could be reused afterwards.

@@

To simulate 120min of seismograms at 17sec resolution, it takes 384 Tesla K20X
GPUs approximately 10 mins to run on Titan. To precisely calculate the sensitivity
kernels, there is forward wavefield saved to the disk and those will used in the
adjoint simulation in the later stage. In the current resolution
and simulation length, the forward wavefield saved is above 1 terabytes
for each earthquake. Given the size the 1480 earthquakes, we have 1.5 petabytes of
wavefield saved to disk using ADIOS\citep{liu2014hello} during about 10 hours
of simulation using 15,360 GPUs.
It is huge volume of data saved to disk in a relatively short period of time, in which case
ADIOS really helped to relief the I/O bottleneck.

@@

The seismic data processing takes the raw observed and synthetic data as input,
and generate s the adjoint sources as the output. As shown
in Fig.\ref{fig:preprocess_workflow}, preprocessing includes:
\begin{enumerate}
  \item Signal processing that remove the instrument response from observed data
    to recover the true ground motion. Both observed and synthetic data will be
    filtered into a certain period band. Then the two horizontal components of the
    seismograms are rotated from North, East to radial and transverse only if the
    StationXML passed the sanity check.
  \item Window selection that picks the time window on a pair of
    observed and synthetic seismograms. Pyflex(python version
    of FLEXWIN\citep{maggi2009automated}), is
    such a tool that automatically generate windows where observed data and
    synthetic data are close enough to make measurements based on user defined
    criteria.
  \item Multi-taper measurement that generates frequency-dependent travel-time
    difference within the windows selected from the last step.
    The output files contains both the measurements and misfit values.
  \item Window cleaning that get rid of bad windows(and associated
    measurements), only keeping good ones based on the statistical analysis of
    measurements.
  \item calculate the adjoint sources with the sorted windows.
  \item For each window and associated measurement, calculate its weightings
    that will be applied at the summation stage and calibrate its contribution
    into the overall misfit function, using the method we mentioned before.
  \item Construct the final adjoint sources by summing the adjoint source from
    different windows and measurement categories.
\end{enumerate}

Even though it is such a computational demanding process as the forward and adjoint
simulation, only taking about $1\%$ CPU hours in the whole workflow, it is crucial
that directly impact the inversion results. It decides which measurement
will be taken into consideration and have a footprint in the final gradient.
We are very careful at this stage, including a lot of checks, such as the Station and
response integrity check, and also filtering out bad measurements to avoid big outliers.
On the other hand, it is very process-demanding, which means it has complex data and
control flows, that every seismograms will go through quite a few of processes
before the final results is generated. We have millions of seismic traces and ten
millions of windows, it is impossible to human-check every trace and measurements.
I will go into more depth about this part in the following sections, discussing our
efforts how to make the data and process safe and robust.

@@

The post-processing takes the frechet derivatives as the input and generate
the model updates as the output. It contains:
\begin{enumerate}
  \item Summation of each kernel for each earthquake to obtain one
    gradient for the overall misfit function. The number of kernel
    files will reduce from 1480 to 1 after summing them up. Also,
    there will source weightings applied at this stage, to balance
    the uneven distribution of earthquakes.
  \item Smoothing the misfit gradient using 3D Gaussian function, which
    serves as a regularization procedure. Instead of using a changing
    smooth radius based upon the "ray density"\citep{bozdaug2016global},
    we used the fixed value in the same iteration, following\cite{zhu2012structure}.
  \item Precondition the smoothed kernels. Following
    \citep{luo2013strategies, zhu2012structure, bozdaug2016global},
    the preconditioner we used is:
    \begin{equation}
      P(\mathbf{x}) = 1 / \int \partial_t^2 \mathbf{s}(\mathbf{x}, t) \cdot \partial_t^2 \mathbf{s}^\dagger(\mathbf{x}, T-t) dt
    \end{equation}
\end{enumerate}

@@

Based upon the misfit gradient obtained from the post-processing stage.
Line search is then used to determine the step length for model update
along the search direction. Forward simulations were run for a subset
of 120 earthquakes with different step length and the one with
the max misfit reduction was picked as the model update step length.

@@

The Adaptable Seismic Data Format, ASDF, serves as a self-contained-and-explained data container while taking full advantages of parallel computing. In our case, one ASDF file contained all the files need in data processing, including the seismic traces, Quakeml files and station response files. The APIs were carefully designed for easy data extraction and parallel implementation. Each trace in the ASDF is associated with one specific Quakeml file and StationXML file, reducing the possibilities of making mistakes.

@@

The inversion went through a few stages, mainly due to adding more data
and measurement categories and fine tuning the weighting strategies.

@@

The major change at stage II is we split the 40-100 sec period band into two: one with
body wave and the other with only surface wave measurements. Since body waves and
surfaces are sensitive to the different parts of the earth, by splitting these two
phases, we have more controls over spatial distribution of the model update.
Table.\ref{table:measurement_category} shows the measurement categories we used starting
from M18 and we will stick to it till M25.

@@

Since surface waves are mostly sensitive to shallow
structures and for shallow structures our model performs quite good
already given current dataset, we don't expect to see large misfit
reduction from in those long period surface wave category.
To better bring out the deep structure, We tune
down the weightings of 40-100 sec surface wave, thus reduce their
contribution in the overall misfit function. That explained why we see the
misfit jump(decrease) from M17 to M18 in Fig.\ref{fig:misfit}(h) - (j) and
increase Fig.\ref{fig:misfit}(b)-(d). We are using the same data but
redistribution the contribution from each category by changing the measurements
weight in the overall misfit function.

@@

The misfit values for 17-40sec and 40-100s body waves are decreasing fast at
the first few iterations, and flattens gradually, which may indicate that we are
approaching the model resolution limits given current dataset.

@@

We added an extra set of 440 earthquakes into our inversion database. Thus, the total
number of earthquakes increased from 1,040 to 1480. We also bring back weightings
of 40-100sec surface wave as stage I since the newly added 440 was mostly located
in poorly covered regions in our previous dataset and those surface wave measurements
will help to better constrain the shallow structures at those regions.

The misfit jumps, from M21 to M22, is mostly due to the redistribution of weights
and new earthquakes being added to the dataset. Optimization method is changed
from conjugate gradient to L-BFGS. Overall misfit values is steadily reducing,
as well as those in each category.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/misfit.pdf}
  \caption{Evolution of phase misfits from GLAD-M15 to GLAD-M25.
  Each color denotes the different stages of inversions, between
  which we make changes to the inversion by either tuning the
  weighting strategies or adding more data.
  (a) Evolution of the total phase misfit. (b) - (j) Misfit in
  each period band and component. (e) - (g) From GLAD-M18,
  we future split the 40-100 sec period band into two
  measurement categories: body(red diamond) waves and surface waves
  (red square).}
  \label{fig:misfit}
\end{figure}

@@

In the last section, we discussed how the misfit evolves in iterations. However, since
we changed our dataset and misfit functions(mainly weightings) between different stages,
it might not be clear that how much did we improve our model. To precisely measure
our model performance, we calculate synthetic data using the same source using M00, M15
and M25. Windows and weightings are also identical across different iterations.

Table.\ref{table:misfit_reduction_M15_M25} shows the misfit reduction from M15 to M25
for 12 categories. For all categories, there is a significant misfit reduction. Among
the three components, transverse have the largest misfit reduction. For the 4 period bands,
40-100s surface waves have the largest misfit reduction which may indicate the 40-100 sec
surface wave might not have good fit between synthetic and observed data compared to
other period band and our inversion is improving it a lot.

Table.\ref{table:misfit_reduction_M00_M25} shows M00 to M25. It is very similar to
Table.\ref{table:misfit_reduction_M15_M25} but with larger values since it also
counts the model improvements from M00 to M15. The best improved category is
the transverse component in 40-100s surface waves, with misfit reduction at
53.8\%. The minimum misfit reduction, on the other hand, is the vertical component
in 17-40s body waves, with 31.1\% misfit reduction.

@@

Another way to evaluate the model performance by checking the distribution
of measurements. Fig.\ref{fig:phase_hist} shows the histograms of phase
measurements. We use identical windows to make measurements using 1,480
earthquakes on the all the three models, our starting model
GLAD-M00(362ANI+CRUST2.0), first generation model GLAD-M15,
and lastest model GLAD-M25. The total number of selected windows is over
18 million. The standard deviations values of phase measurements has
significant improvements on all the categories over iterations.
The means values, has also improved especially in the long period categories,
such as the 40-100 surface waves and 90-250 all waves. In some categories,
you may see slight shift of mean values, mainly due to the trade off among
different components and period bands. To summarize, the improvements in
the phase measurements histogram is consistent as we observed in the
misfit reduction.

@@

The other prominent feature of our models is that we can also constrain $V_p$ as an
independent model parameters. Fig.\ref{fig:global-vp} shows our isotropic $V_p$
structures at several depths. As comparisons, we also showed two more global
$V_p$ models, LLNL-G3Dv3\citep{simmons2012llnl}} and GAP-P4\citep{fukao2013subducted}.

@@

For upper mantle, the LLNL-G3Dv3 shows the largest variations. For example,
the 500km, the LLNL model shows very strong slow velocity anomalies at
Hawaii and other pacific hotspots, such as Samoa and Tahiti.
For middle mantle at about 1000km, all them models all shows a few major
subduction zones, such as the South America, Fiji and Sumatra regions.
At lower mantle, such as 2000km and 2800km, LLNL-G3Dv3 and GAP-P4 tends to
fade out, even though the patterns are very similar among them, such as
the LLSVP regions.

** Note: add spherical harmonic analysis later **

@@

\subsubsection{North America}
The deployment of USArray has dramatically increased the data overage on the
continental United States. In this section, we compare our GLAD-M25 with
together with two global models, S362ANI+M, S40RTS, and two regional models,
US22\citep{zhu2017radial}, SL2013NA\citep{schaeffer2014imaging}.

US22 is radially anisotropic model build based on adjoint tomography using
USArray data from 180 regional earthquakes using 15-50sec short-period
body waves and 25-150sec long-period surface waves.
SL2013NA is an upper-mantle global shear wave speed model using multimode
surface waveform tomography and primarily focused on the north American
region.

The North American region is featured by a large, high-velocity cratonic
lithosphere, bounded by the Rocky Mountain Front to the west and continental
margin to the east. For both sides the lithosphere has been deformed by
low velocity structures thus form a sharp boundary.

At 100km, our model shows a much sharpened boundary between the carton
and surrounding low velocity structures, especially compared with
two global models, S362ANI+M and S40RTS, yet stay very similar to
two regional models, US22 and SL2014NA. At such depth,
we are observed a very strong expression of the Snake River Plain(
with Yellowstone) and Raton hotspot, with their depth extending to 200km.
The very high-velocity craton extend to 200km, with its very core part to
300 km, similar to US22 while SL2013NA shows it almost disappeared at such depth.
Again, the Wyoming craton and Medicine Hat Block shows a very strong lithosphere
root from the very shallow depth to 400km, which is also similar to US22 but
hard to observe in SL2013NA. Similar case is the craton at Yavapai and
Mazatzal Blocks, which define the southern bound of the large craton.

At such depth, we also observed high velocity
features within the Gulf of Mexico, which coincides spatially with the deepest
bathymetry and corresponds to a portion of ancient oceanic
Lithosphere\citep{muller2008}. The remnant of Juan de Fuca subduction
plates, with its high velocity footprint, is presented in our model at
300km and 400km, though the SL2013NA shows a much stronger anomaly.

At the North, our models shows a extended high velocity craton beneath the 
Greenland, though clearly separated with the North America Craton,
bounded at the relatively low velocity structure at Barffin Bay and Labrador
Sea, with North American Craton followed nicely with the coastline. In the
north west region, our models shows the most clear picture of
Aleutian subduction zones, with it high velocity feature down to 400km.
In the northeastern part, the Iceland hotspot has been greatly enhanced,
especially at 300km and 400km compared to S362ANI+M.

In the Easter part, our models shows a high-velocity structure at the western
Atlantic, starting from the tip of Florida to the New York,
separated from the major part the north America craton,
though it stays very thin that the depth only extend to above 200km.
This feature is associated with some of the oldest oceanic crust and lithosphere
within the Atlantic\citep{muller2008, schaeffer2014imaging}.
All the models is showing a high velocity anomaly at Nova Scotia and Newfoundland
with its high-velocity core to 300km.
Bermuda hotspot is nicely preserved from our starting
model, yet get sharpened and stronger at 200km and 300km. Besides, the
Caribbean suduction, which is relatively young subduction systems, is clearly
observable in our model, starting from 200km to 300km.

@@

\subsubsection{Europe}

...to be done...
Fig.\ref{fig:europe-vs} shows the isotropic $V_s$ structures at Europe. As
comparisons, we also showed S362ANI+M, S40RTS and one regional model EU60
\citep{zhu2015seismic}. Compared to two global models, it is clear that
our model is much sharper at few locations. At 100 and 200km, Eifel
hotspot is clearly seen in our models, as well as in EU60. At 400km
and 600km, the subduction is very consistent between GLAD-M25 and EU60,
with the curved shape at the lower bottom of the figure,
while S362ANI+M shows very smooth shape and S40RTS shows much smaller 
perturbation values. Besides, at 400km, GLAD-M25 is enhancing Iceland
compared to S362ANI+M.

@@

...to be done...
We presented our model GLAD-M25, with S362ANI+M, S40RTS and EARA2014.
At 100km depth, our model shows sharpening at the Himalayas subductions
zones. At 250km, our models shows the subduction in the east part.
At 400km, our models shows the subduction at sumartra.

@@

\subsubsection{South America}
...to be done...

@@

\subsection{Held-out Dataset Test}
Following the approach by \cite{tape2009adjoint}, \cite{bozdaug2016global},
unseen events are used to check the performance of our
model. We revisited the Global CMT catalogue,
picked all the events whose magnitude is between 6.3 and 7.0, which hasn't
been used in our inversion. The reason why we set the lower bound to 6.3
instead of 5.5, is those earthquakes have good enough data and generates
more measurements at the global scale.
We calibrated the scalar moment and origin time in the GLAD-M25, and there
are 360 events left that has enough windows. Fig.\ref{fig:events_360} shows
the distribution of 360 earthquake that we used as held-out dataset

@@

First, we evaluate the model performance by misfit reduction. Same windows and
weightings are used across models. Table.\ref{table:misfit_reduction_M00_M25_360}
shows the misfit reduction from M15 to M25 and Table.\ref{table:misfit_reduction_M00_M25_360}
is from M00 to M25. It again shows misfit reduction in all categories in
our held-out dataset, similar to what we observed from the inversion dataset.
The reduction are slightly smaller, partly due to the reason that we didn't
do waveform-type inversion as we mentioned in the source inversion stage but
only did a grid search to calibrate the scalar moment and origin time.
